Manuscript figures

### 2024-03-06 04_manuscript_plots.Rmd
type  --> class

```{r model-type, fig.width=10}
top10_temp <- 
  temperature_skill |> 
  group_by(model_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ungroup() |>
  filter(mean_crps > 0) |>
  left_join(model_meta) |> 
  mutate(manuscript_name = fct_reorder(manuscript_name, mean_crps),
         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no'), 
         uses_AT = str_detect(NOAA_var, 'air_temperature')) |> 
  ggplot(aes(x=mean_crps, y = manuscript_name, fill = model_type)) +    # Different pattern for each group
  geom_bar_pattern(aes(pattern = as.factor(uses_AT)),
                   stat = "identity",
                   alpha = 0.8, 
                   pattern_fill = "black",
                   pattern_colour = "black", 
                   pattern_spacing = 0.04,
                   pattern_alpha = 1,
                   colour = 'black') + 
  scale_y_reordered() +
  labs(y='Model', x = 'mean relative skill (°C)') +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = cols_modeltype) +
  scale_pattern_manual(values=c('none', 'stripe')) +
  scale_pattern_type_manual(values=c(NA, NA)) +
  theme_bw() + 
  guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02), 
                                title = 'Uses air temperature as a covariate?'),
         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))

@@

  geom_bar_pattern(aes(pattern = as.factor(uses_AT)),
                   stat = "identity",
                   alpha = 0.8, 
                   colour = 'black',
                   pattern_fill = "black",
                   pattern_colour = "black", width = 0.8,
                   pattern_spacing = 0.03,
                   pattern_alpha = 1) + 
  scale_y_reordered() +
  labs(y='Model', x = 'mean relative skill (°C)') +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = cols_modeltype) +
  scale_pattern_manual(values=c('none', 'stripe'), name = 'Uses air temperature as a covariate?') +
  scale_pattern_type_manual(values=c(NA, NA), name = 'Uses air temperature as a covariate?') +
  theme_bw() + 
  guides(pattern = guide_legend(override.aes = list(fill = "white"), title = 'Uses air temperature as a covariate?'),
         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))

@@

bias <- temperature_scores |> 
  mutate(horizon = as_date(datetime) - as_date(reference_datetime),
         abs_bias = abs(mean - observation)) |> 
  filter(model_id %in% top_mods$model_id & model_id != 'persistenceRW', 
         horizon > 0, 
         horizon != 15,
         horizon <= 30) |>
  group_by(model_id, horizon) |> 
  summarise(mean_abs_bias = mean(abs_bias, na.rm = T)) |> 
  left_join(select(model_meta, model_id, manuscript_name)) |> 
  mutate(manuscript_name = factor(manuscript_name, levels = levels(model_id_colours$manuscript_name))) |> 
  ggplot(aes(x=horizon, 
             y = mean_abs_bias, colour = manuscript_name)) +
  geom_line()+
  labs(y='mean absolute bias (°C)', x = 'Horizon (days)') +
  scale_colour_manual(name = 'Model',values = model_id_colours$value) +
  theme_bw() +
  coord_cartesian(ylim = c(0, 2.5))

@@

A <- temperature_skill |> 
  left_join(model_meta, by = 'model_id') |>
  filter(model_type != 'null') |> 
  group_by(model_type, model_id, site_id) |> 
  summarise(mean_skill = mean(skill_crps, na.rm = T)) |>  
  ggplot(aes(x = mean_skill, 
             y = site_id, 
             colour = model_type)) +
  geom_vline(xintercept = 0, linetype = 'solid') +
  geom_label(data = n_model_type, aes(label = paste0('n = ', n)))+ 
  geom_boxplot() +
  facet_wrap(~model_type, scales = 'free_y',nrow = 2) +
  labs(y='Site', x = 'mean relative skill (°C)') +
  scale_colour_manual(values = cols_modeltype, name = 'Model Type') +
  theme_bw()

B <- temperature_skill |> 
  filter(model_id %in% top_mods$model_id) |> 
  group_by(site_id, horizon) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  mutate(site_id = factor(site_id, levels = lake_sites)) |> 
  ggplot(aes(x=horizon, 
             y = mean_crps, colour = site_id)) +
  geom_hline(yintercept = 0, colour = 'grey40') + 
  geom_line()  +  
  labs(y='mean relative skill (°C)', x = 'Horizon (days)') +
  scale_colour_brewer(palette = 'Dark2')+
  theme_bw()

>>>>

```{r model-type, fig.width=10}
top10_temp <- 
  temperature_skill |> 
  group_by(model_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ungroup() |>
  filter(mean_crps > 0) |>
  left_join(model_meta) |> 
  mutate(manuscript_name = fct_reorder(manuscript_name, mean_crps),
         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no'), 
         uses_AT = str_detect(NOAA_var, 'air_temperature')) |> 
  ggplot(aes(x=mean_crps, y = manuscript_name, fill = model_type)) +    # Different pattern for each group
  geom_bar_pattern(aes(pattern = as.factor(uses_AT)),
                   stat = "identity",
                   alpha = 0.8, 
                   pattern_fill = "black",
                   pattern_colour = "black", 
                   pattern_spacing = 0.04,
                   pattern_alpha = 1,
                   colour = 'black') + 
  scale_y_reordered() +
  labs(y='Model', x = 'mean relative skill (°C)') +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = cols_modeltype) +
  scale_pattern_manual(values=c('none', 'stripe')) +
  scale_pattern_type_manual(values=c(NA, NA)) +
  theme_bw() + 
  guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02), 
                                title = 'Uses air temperature as a covariate?'),
         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model class'))

@@

  geom_bar_pattern(aes(pattern = as.factor(uses_AT)),
                   stat = "identity",
                   alpha = 0.8, 
                   colour = 'black',
                   pattern_fill = "black",
                   pattern_colour = "black", width = 0.8,
                   pattern_spacing = 0.03,
                   pattern_alpha = 1) + 
  scale_y_reordered() +
  labs(y='Model', x = 'mean relative skill (°C)') +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = cols_modeltype) +
  scale_pattern_manual(values=c('none', 'stripe'), name = 'Uses air temperature as a covariate?') +
  scale_pattern_type_manual(values=c(NA, NA), name = 'Uses air temperature as a covariate?') +
  theme_bw() + 
  guides(pattern = guide_legend(override.aes = list(fill = "white"), title = 'Uses air temperature as a covariate?'),
         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model class'))

@@

bias <- temperature_scores |> 
  mutate(horizon = as_date(datetime) - as_date(reference_datetime),
         abs_bias = abs(mean - observation)) |> 
  filter(model_id %in% top_mods$model_id & model_id != 'persistenceRW', 
         horizon > 0, 
         horizon != 15,
         horizon <= 30) |>
  group_by(model_id, horizon) |> 
  summarise(mean_abs_bias = mean(abs_bias, na.rm = T)) |> 
  left_join(select(model_meta, model_id, manuscript_name)) |> 
  mutate(manuscript_name = factor(manuscript_name, levels = levels(model_id_colours$manuscript_name))) |> 
  ggplot(aes(x=horizon, 
             y = mean_abs_bias, colour = manuscript_name)) +
  geom_line()+
  labs(y='mean absolute bias (°C)', x = 'Horizon (days)') +
  scale_colour_manual(name = 'Model',values = model_id_colours$value) +
  theme_bw() +
  coord_cartesian(ylim = c(0, 2))

@@

A <- temperature_skill |> 
  left_join(model_meta, by = 'model_id') |>
  filter(model_type != 'null') |> 
  group_by(model_type, model_id, site_id) |> 
  summarise(mean_skill = mean(skill_crps, na.rm = T)) |>  
  ggplot(aes(x = mean_skill, 
             y = site_id, 
             colour = model_type)) +
  geom_vline(xintercept = 0, linetype = 'solid') +
  geom_text(data = n_model_type, aes(label = paste0('n = ', n)))+ 
  geom_boxplot() +
  facet_wrap(~model_type, scales = 'free_y',nrow = 2) +
  labs(y='Site', x = 'mean relative skill (°C)') +
  scale_colour_manual(values = cols_modeltype, name = 'Model class') +
  theme_bw()

B <- temperature_skill |> 
  filter(model_id %in% top_mods$model_id) |> 
  group_by(site_id, horizon) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  mutate(site_id = factor(site_id, levels = lake_sites)) |> 
  ggplot(aes(x=horizon, 
             y = mean_crps, colour = site_id)) +
  geom_hline(yintercept = 0, colour = 'grey40') + 
  geom_line()  +  
  labs(y='mean relative skill (°C)', x = 'Horizon (days)') +
  scale_colour_brewer(palette = 'Dark2', name = 'Site')+
  theme_bw()
